---
title: "Kinship Visualization"
author: "K. W. Bridges"
date: '2022-04-25'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r libraries}
library(kinship2)   ## Generate pedigree diagrams
library(tidyverse)  ## Data wrangline
library(flextable)  ## Make good tables
library(ftExtra)    ## Allows markdown code inside a flextable
library(R.devices)  ## Capture plots (whew!! This is a big "trick.")
library(ggrepel)    ## Used to caption legends
library(patchwork)  ## Assemble graphic elements (e.g., pedigree + legend)
library(ggimage)    ## Read PNG images
library(magick)     ## Useful image manipulation

```

## Pedigree Diagrams Using `kinship2`

The basic purpose of a pedigree diagram is to show the relationships between relatives. This is a straightforward task for simple sets of relationships.

The R package `kinship2` is the key. It uses a basic set of information to draw the pedigree diagram. The `kinship2` package does much more than produce the relationship visualization. Here, however, we're focusing just on the diagrams that are created using the package.

For more information on `kinship2`: https://cran.r-project.org/package=kinship2

## Basic data structures

The following five columns are, essentially, the minimal data structure. The names of the fields are pretty standard for a pedigree chart.

* ID: Use a person's name
* Father: Person's name
* Mother: Person's name
* Sex: Gender as `male`, `female` or `NA` or coded as (1=male, 2=female, 0=NA)
* Status: Code as living=0 or dead=1 (dead is a strikethrough diagonally across the symbol)

### A few data rules

Most uses of `kinship2` will involve a straightforward data file. 

* Each individual is a row in the data structure.

* The order of the rows is not important.

* When an individual's parents are not known, code each parent as NA (or 0).

* Adding a spouse when there are no children requires special handling (shown later).

\newpage
```{r basic}

## Basic data
data <- read_csv(col_names = TRUE, file=
    "ID,    dad,  mom, sex,    status 
     One,   NA,   NA,  male,   1     
     Two,   NA,   NA,  female, 1      
     Three, One,  Two, male,   0     
     Four,  One,  Two, male,   1      
     Five,  One,  Two, male,   0      
     Six,   NA,   NA,  female, 0      
     Seven, Five, Six, female, 0")

## Print a table
flextable(data) %>% 
  autofit()

## Generate the pedigree data structure (without no-children couples)
ped <- pedigree(id=data$ID, 
                dadid=data$dad, 
                momid=data$mom, 
                sex=data$sex,
                status = data$status)

## Plot the pedigree chart
plot(ped)

## Save the plot as a PNG file
devEval("png", 
        path = "P://Hot/kinship", 
        filename="pedigree_basic.png", 
        width=7.5, units="in", res=300,
        {
  plot(ped)
  })

```

\newpage

## Relationships Without Children

The basic pedigree diagram is the starting point for this example. Here, the spouses of siblings `Three` and `Four` are added. One of these (`Three`), like sibling `Seven`, has a child. The other (`Four`) doesn't have a child. Handling the "childless" sibling requires a different data structure.

Each couple without children needs to have a special set of relationship information. We'll call this `pairs` here.

First, make sure that all individuals are included in the basic data. Each person in the pedigree must have a data row.

Then, add a new data structure that has three fields (`id1`, `id2`, `code`). _Use these field names_. The code field value of `4` indicates that this is a spouse. 

The set of relationship data is added to the basic pedigree data in the `pedigree` function.

```{r pairs}

## Basic data
data <- read_csv(col_names = TRUE, file=
    "ID,    dad,   mom,   sex,    status 
     One,   NA,    NA,    male,   1     
     Two,   NA,    NA,    female, 1      
     Three, One,   Two,   male,   0     
     Four,  One,   Two,   male,   1      
     Five,  One,   Two,   male,   0      
     Six,   NA,    NA,    female, 0      
     Seven, Five,  Six,   female, 0
     Eight, NA,    NA,    female, 0
     Nine,  Three, Eight, male,   0
     Ten,   NA,    NA,    female, 0")

## Relationships without children
pairs <- read_csv(col_names = TRUE, file=
                     "id1,  id2,  code
                      Four, Ten,  4")

## Generate the pedigree data structure (with no-children couples)
ped <- pedigree(id=data$ID, 
                dadid=data$dad, 
                momid=data$mom, 
                sex=data$sex,
                status = data$status,
                relation = pairs)     ## This adds the no-children spouses

## Plot the pedigree chart
plot(ped) 

```

\newpage
## Long Names

A real set of data will, most likely, use peoples' names. The result might be crowding of the pedigree diagram. One solution is to split the names onto multiple lines.

Splitting a name uses the code for `\\n` to create a new line. This can't always be coded directly into the data as it may affect the reading of the data lines. Instead, a function is used to replace each blank space with the new line code.

## Scaling the plot

You can control the size of the symbols and identifying labels.

* `symbolsize`: Use this to make the symbol larger (>1) or smaller (<1).
* `cex`: The text is scaled with `cex`. Values <1 make the text smaller.

```{r names}

## Basic data
data <- read_csv(col_names = TRUE, file=
    "ID,                 dad,           mom,         sex,    status 
     William Jones,      NA,            NA,          male,   1     
     Mary White,         NA,            NA,          female, 1      
     Frank Jones,        William Jones, Mary White,  male,   0     
     Mike Jones,         William Jones, Mary White,  male,   1   
     Peter Jones,        William Jones, Mary White,  male,   0     
     Carol Smith,        NA,            NA,          female, 0      
     LaVerne Jones,      Peter Jones,   Carol Smith, female, 0
     Karen Brown,        NA,            NA,          female, 0
     Samuel Jones-Brown, Frank Jones,   Karen Brown, male,   0
     Susan Hanford,      NA,            NA,          female, 0")

## Relationships without children
pairs <- read_csv(col_names = TRUE, file=
                     "id1,        id2,            code
                      Mike Jones, Susan Hanford,  4")

## Put in a line break to stack the names
data$ID   <- gsub(" ","\n",data$ID)
data$dad  <- gsub(" ","\n",data$dad)
data$mom  <- gsub(" ","\n",data$mom)
pairs$id1 <- gsub(" ","\n",pairs$id1)
pairs$id2 <- gsub(" ","\n",pairs$id2)

## Generate the pedigree data structure (with no-children couples)
ped <- pedigree(id=data$ID, 
                dadid=data$dad, 
                momid=data$mom, 
                sex=data$sex,
                status = data$status,
                relation = pairs)

## Plot the pedigree chart
plot(ped,
     cex = 0.8,
     symbolsize = 0.7) 

```

\newpage

## Generalizing the Use of a Pedigree

The previous pedigree diagrams show basic relationships (symbols connected with lines), each person's gender (square vs circle) and whether a person is dead (diagonal strike through) or alive.

But *much* more information can be conveyed in the diagram about each individual.

The `kinship2` package permits many "customizations." This is best introduced with an example.

Note that the additional visualization features are data driven. You re-purpose existing data fields and add additional places to store information. This information gets integrated into `kinship2` using existing linkages.

### The Example Scenario

The family (using the same names as the previous examples) has three generations. The grandparents are dead and the inheritance is being divided. Grandpa William was a staunch conservative (red) while his loyal wife Mary has been a closet liberal (blue). Grandpa's will shows only some of their family are going to get any inheritance. Son Frank (a teacher) and his wife Karen (a lawyer) have a liberal son Samuel who is following his father's profession. This branch of the family is way too liberal for Grandpa William so they have been cut out of any inheritance. Son Mike and his wife Susan are both conservative lawyers. Grandpa is OK with that. The third son, Peter, has taken over the farm, along with his wife Carol. They have a daughter LaVerne who, much to the dismay of her conservative parents, is a radical Green-party follower and is showing ambiguity about choosing farming as a career. Grandpa disapproves and specifically cuts her out of any inheritance. 

Note that the symbolism used in the story is captured in the pedigree diagram. The problem is that a legend is needed to explain some of the symbolism. That will be treated next.

```{r max}

## Family relationship
family <- read_csv(col_names = TRUE, file=
    "ID,                 dad,           mom
     William Jones,      NA,            NA     
     Mary White,         NA,            NA     
     Frank Jones,        William Jones, Mary White
     Mike Jones,         William Jones, Mary White
     Peter Jones,        William Jones, Mary White
     Carol Smith,        NA,            NA  
     LaVerne Jones,      Peter Jones,   Carol Smith
     Karen Brown,        NA,            NA
     Samuel Jones-Brown, Frank Jones,   Karen Brown
     Susan Hanford,      NA,            NA")

## Info about the family members
info <- read_csv(col_names = TRUE, file=
    "ID,                sex, inherit, party, teacher, lawyer, farmer, housewife
     William Jones,      1,      0,   red,      0,       0,      1,      0     
     Mary White,         2,      0,   blue,     0,       0,      0,      1        
     Frank Jones,        1,      1,   blue,     1,       0,      0,      0         
     Mike Jones,         1,      0,   red,      0,       1,      0,      0     
     Peter Jones,        1,      0,   red,      0,       0,      1,      0       
     Carol Smith,        2,      0,   red,      0,       0,      0,      1         
     LaVerne Jones,      2,      1,   seagreen, 0,       0,      NA,     0    
     Karen Brown,        2,      1,   purple,   0,       1,      0,      0   
     Samuel Jones-Brown, 1,      1,   blue,     1,       0,      0,      0   
     Susan Hanford,      2,      0,   red,      0,       1,      0,      0     ")

## Merge the data
## This results in sorting by ID
data <- merge(family, info, by = "ID")

## Sort the affected list
aff2 <- info %>% 
  arrange(ID) 

## Relationships without children
pairs <- read_csv(col_names = TRUE, file=
                     "id1,        id2,            code
                      Mike Jones, Susan Hanford,  4")

## Put in a line break to stack the names
data$ID   <- gsub(" ","\n",data$ID)
data$dad  <- gsub(" ","\n",data$dad)
data$mom  <- gsub(" ","\n",data$mom)
pairs$id1 <- gsub(" ","\n",pairs$id1)
pairs$id2 <- gsub(" ","\n",pairs$id2)

## Generate the pedigree data structure (with no-children couples)
ped <- pedigree(id=data$ID, 
                dadid=data$dad, 
                momid=data$mom, 
                sex=data$sex,
                status = data$inherit,
                relation = pairs)

## Plot the pedigree chart
plot(ped,
     cex = 0.8,
     symbolsize = 1.8,
     col = data$party,
     density = c(-1, -1, -1, -1), ## makes all of the sections solid!
     affected = cbind(aff2$teacher,aff2$lawyer,aff2$farmer,aff2$housewife))

## Save plot as a PNG file
devEval("png", 
        path = "P://Hot/kinship", 
        filename="pedigree_jones.png", 
        width=7.5, units="in", res=300,
        {
  plot(ped,
     cex = 0.8,
     symbolsize = 1.8,
     col = data$party,
     density = c(-1, -1, -1, -1), ## makes all of the sections solid!
     affected = cbind(aff2$teacher,aff2$lawyer,aff2$farmer,aff2$housewife))
  })

```

\newpage
## Creating a Key

A key is needed to explain the color and segment conventions. 

The `kinship2` package has a built-in legend capability, but this isn't very effective for the more general use of the package. Instead, custom diagrams are created for the purpose of identifying the symbolic notation.

The symbol segmentation is handled with a pie-shaped diagram with labels identifying the role of each segment. The color of the symbol use colored boxes with identifying names inside.

The construction of the pie and box diagrams is a bit complicated. However, the use of this code is very straightforward as these are data-driven procedures. The only adjustment is likely the size of the label text.

### Symbol segments

The symbol, whether it is a circle or square, is filled with from one to four segments. Each segment can be empty, filled or have a small question mark.

The order of the segments is the same whether they are inside a square or circle. Therefore, the key only uses a circle (i.e., pie-shape) to show the value of each segment. (It is way easier to produce!)

The symbols can come in many different colors, but that isn't related specifically to the value associated with the segments as all the segments for an individual are shown in the same color. This means that the legend for the symbol segments can be a neutral color, like gray or black.

There will be some scaling adjustment needed, especially with the text, when the legend is combined with the pedigree diagram.

```{r segments}

## Each of the affected variables needs to have a name assigned. 
## The list below can be from one to four items.

## Read information to create the legend (as segments of a pie)
## The size should be the same value for each segment
pie <- read_csv(col_names = TRUE, file = 
                  "name,     size, fill
                   teacher,     1, black
                   lawyer,      1, black
                   farmer,      1, black
                   housewife,   1, black")

## Calculate the position of each label
pie <- pie %>% 
  mutate(csum = rev(cumsum(rev(size))), 
         pos = size/2 + lead(csum, 1),
         pos = if_else(is.na(pos), size/2, pos))

## Generate the legend
legend <- ggplot(pie, aes(x="", y=size), fill=fill) +
  geom_bar(stat="identity", 
           width=1,           ## width of lines inside the pie
           color = "gray",    ## color of lines inside the pie
           show.legend = FALSE) +
  coord_polar("y", start=0) +  ## pie shape
  geom_label_repel(data = pie,
                   aes(y = pos, label = name),
                       size = 3.0,          ## label text size
                       segment.color = "gray", ## connector line color
                       segment.size  = 0.8,    ## connector line width
                       nudge_x = 1, 
                       show.legend = FALSE, 
                       fill = "white") +    ## label background
  ggtitle("Key") +
  theme_void()

legend

```

\newpage
### Symbol colors

The symbol for each individual in a pedigree can be a unique color. The entire symbol (i.e., all the segments and the surrounding line) are shown in a single color. You can't make each segment a different color.

The color matching to a name is done here with colored boxes, one for each color. Inside each box is the name associated with the color.

As with the symbol segment key, the procedure looks complicated. But this, too, is simply data driven. As with symbol segments, there may need to be an adjustment of the text size.

```{r strike}

## Make boxes with adjacent labels (fill with colors)
boxes <- read_csv(col_names = TRUE, file =
                    "name,         color
                     conservative, red
                     independent,  purple
                     liberal,      blue
                     green,        seagreen")

## Calculate the box corners
col_number <- length(boxes$name)
x1 <- rep(1, times=col_number)
x2 <- rep(2, times=col_number)
y1max <- (2*col_number) -1
y1 <- seq(1, y1max, by=2)
y2max <- (2*col_number)
y2 <- seq(2, y2max, by=2)

## Put data into a data structure
d=data.frame(x1, x2, y1, y2,
             r=boxes$name,
             t=boxes$color)

## Generate the plot
box <- ggplot()+
geom_rect(data=d, 
          mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2),
          fill=d$t, 
          color="black", 
          alpha=1.0,
          show.legend = FALSE) +
geom_text(data=d, 
          aes(x=x1+(x2-x1)/2, y=y1+(y2-y1)/2, label=r), 
          size=4,
          color="white") +
  theme_void()
  
## Plot the boxes
box

```

\newpage
### Assembling the pedigree and key

The `patchwork` package provides some very flexible ways to format the pieces of the pedigree for use as a final diagram.

Some adjustment is required to get a balance in the size of each pedigree element. There are likely adjustments needed in the size of the labels.

The resulting diagram shows the type of complexity that can be achieved using the `kinship2` package.

### Saving the pedigree diagram

The `R.devices` package was used earlier to save the pedigree diagram when it was plotted using the `kinship2` plot function. That introduced some (necessary) complexity.

Here, it is possible to use the `ggsave` function. This is much more straightforward and, likely, less error prone.


```{r assembly}

## Use patchwork to arrange the three components

## Retrieve the pedigree image
p <- image_read2("P://Hot/kinship/pedigree_jones.png")

## Provide a patchwork layout guide
layout <- "
AAAAAAA#BB
AAAAAAA#BB
AAAAAAA#CC
"

## Composite the pedigree and the key
p_out <- image_ggplot(p) + legend + box + plot_layout(design = layout)

## Generate the output
p_out

## Save the output as a PNG file
ggsave(filename="P:/Hot/kinship/pedigree_jones_all.png",
       plot=p_out,
       width=7.5, 
       height=6, 
       units = "in")

```

\newpage
## Calling Attention to Specific Individuals

The previous example was complicated. Most pedigrees don't need to use very many categories of information. Highlighting a few specific individuals is adequate.

Here, we focus on modify the standard symbolism with two, data driven additions: color and solid fill.

`color`: This will color the line surrounding the symbol. It is also be the color of the filled pattern inside the symbol, if the `affected` value indicates filling.

`affected`: This controls pattern inside the symbol. The default is empty. Note that in the following example, there are two values for `affected`. One is for the left side of the symbol, the other for the right side. You can use only one of these, if desired. Apparently, up to four columns of `affected` can be used. The following values can be used:

* 0 = empty
* 1 = filled
* NA = "?"

```{r color}
## Basic data
data <- read_csv(col_names = TRUE, file=
    "ID,                 dad,           mom,         sex,    status, color,      affected 
     William Jones,      NA,            NA,          male,   1,      black,      NA     
     Mary Jones,         NA,            NA,          female, 1,      blue,       1      
     Frank Jones,        William Jones, Mary Jones,  male,   0,      black,      0     
     John Jones,         William Jones, Mary Jones,  male,   1,      black,      0   
     Peter Jones,        William Jones, Mary Jones,  male,   0,      olivedrab,  1     
     Carol Jones,        NA,            NA,          female, 0,      blue,       1      
     LaVerne Jones,      Peter Jones,   Carol Jones, female, 0,      black,      0
     Karen Brown,        NA,            NA,          female, 0,      black,      0
     Samuel Jones-Brown, Frank Jones,   Karen Brown, male,   0,      olivedrab,  1
     Susan Hanford,      NA,            NA,          female, 0,      black,      NA")

## Relationships without children
pairs <- read_csv(col_names = TRUE, file=
                     "id1,        id2,            code
                      John Jones, Susan Hanford,  4")

## Put in a line break to stack the names
data$ID   <- gsub(" ","\n",data$ID)
data$dad  <- gsub(" ","\n",data$dad)
data$mom  <- gsub(" ","\n",data$mom)
pairs$id1 <- gsub(" ","\n",pairs$id1)
pairs$id2 <- gsub(" ","\n",pairs$id2)

## Generate the pedigree data structure (with no-children couples)
ped <- pedigree(id=data$ID, 
                dadid=data$dad, 
                momid=data$mom, 
                sex=data$sex,
                status = data$status,
                relation = pairs)

## Plot the pedigree chart
plot(ped,
     cex = 0.8,
     symbolsize = 1.2,
     col = data$color,
     affected = data$affected) 


```


\newpage
## More About Symbolism

The default values for the symbols is to have different _shade patterns_ for each segment. The examples shown so far have used solid segments.

Here, we'll look at the shade pattern alternatives. 

The following pedigree diagram uses a mixture of values for the `affected` parameter. Colors can vary, too. The colors in this example are all set to "black" as shading is generally used when printing will be done in black and white.

Note that the shade comes from two properties, the density of the lines and the angle of the lines. These properties are controlled independently.

Each segment has a set combination of line density and angle. This can't differ between individuals.

```{r shade1}

data <- read_csv(col_names = TRUE, file=
    "ID,    dad,  mom, sex,    status, color,  aff1, aff2, aff3, aff4
     One,   NA,   NA,  male,   1,      black,   0,    NA,   1,    1
     Two,   NA,   NA,  female, 1,      black,   NA,   NA,   1,    0
     Three, One,  Two, male,   0,      black,   0,    1,    0,    1
     Four,  One,  Two, male,   0,      black,   0,    1,    0,    0
     Five,  One,  Two, male,   0,      black,   1,    1,    NA,   1
     Six,   NA,   NA,  female, 0,      black,   NA,   0,    1,    NA
     Seven, Five, Six, female, 0,      black,   1,    0,    NA,   NA")

## Print a table
flextable(data) %>% 
  autofit()

## Generate the pedigree data structure (with no-children couples)
ped <- pedigree(id=data$ID, 
                dadid=data$dad, 
                momid=data$mom, 
                sex=data$sex,
                status = data$status)

## Plot the pedigree chart using default shade values
plot(ped, 
     cex=0.8, 
     col=data$color, 
     symbolsize = 2.5, 
     affected =cbind(data$aff1, data$aff2, data$aff3, data$aff4))   

```

\newpage
### Controlling the shade pattern

The two elements making up the shade pattern are shown below.

These tables have the default density values (used, successively for the four affected variables)

Line spacing (default)

| value | line spacing |
|---:|---:|
| -1 | solid |
| 35 | medium density |
| 65 | densely spaced |
| 20 | widely spaced |

Line angle (default)

| value | line pattern |
|---:|---:|
| 90 | vertical |
| 65 | steeply inclined |
| 40 | nearly diagonal |
| 0 | horizontal |

These values are changed with a parameter in the `kinship2` plot parameter. This is seen in the next example where the previous data are used. Only the plotting parameters are changed.

```{r shade2}

## Plot the pedigree chart using modified shade values
plot(ped, 
     cex        = 0.8, 
     col        = data$color, 
     symbolsize = 2.5,
     density    = c(30, 40, 40, 30),  ## line spacing
     angle      = c(90, 45, -45, 0),  ## line angle
     affected   = cbind(data$aff1, data$aff2, data$aff3, data$aff4))  

```





```{r }


```

\newpage
## Parting Thoughts

The `kinship2` package is likely very useful to people who handle complex problems involving pedigree information. It is nice that a non-specialist can use this package for simple problems without too much difficulty. 

Having a pedigree display program in R is important. This allows an individual to stay within the R-universe and to integrate pedigree information with other tabular and visualization activities. `kinship2` fills most of this requirement.

### Limitations

It's not obvious how to modify the width of the lines in the plot. Therefore, things like all the connector lines are very thin.

The plotting should be more compatible with the standards used in `ggplot2`. That would be a very handy simplification. A work-around was found, so this is not really a limitation.

### Gratitude

The authors of `kinship2` and the contributors to the development of this package need to be thanked for their contribution.

* Jason P. Sinnwell
* Terry M. Therneau
* Elizabeth J. Atkinson
* Daniel J. Schaid
* Carly Mester 

## Software Versions

* `kinship2` (1.8.5)
* `tidyverse` (1.3.1)
* `flextable` (0.6.7)
* `ftExtra` (0.2.0)
* `R.devices` (2.17.0)
* `ggrepel` (0.9.1)
* `patchwork` (1.1.1)
* `ggimage` (0.2.8)
* `magick` (2.7.2)

The CRAN page for `kinship2` is found at https://cran.r-project.org/package=kinship2

\newpage

## A Wrapper Experiment

Sometimes, you just want to do the minimal work to get a pedigree diagram. What can you do?

If you name the fields in your data with standard names, you can create some good, basic results.

Here are the field names (case sensitive) for basic relationship data:

* ID     (required)
* dad    (required)
* mom    (required)
* sex    (required)
* status (optional, if not used, assumed value is 0=no strikethrough)
* color  (optional, if not used, assumed value is black")
* fill   (optional, if not used, assumed value is 0=no fill)

Pairs data, as before, requires exact field names. Here, the `code` field is optional.


```{r wrap}

ez_ped <- function(data, pairs=NA, fold=TRUE, symbolsize=1.8, cex=0.8){
  
## Checks the data fields
`%!in%` <- Negate(`%in%`) ## create the operator  
if("color"  %!in% colnames(data)) {data$color  <- "black"}
if("fill"   %!in% colnames(data)) {data$fill   <- 0}
if("status" %!in% colnames(data)) {data$status <- 0}


## Put in a line break to stack the names
if(isTRUE(fold)){  
   data$ID   <- gsub(" ","\n",data$ID)
   data$dad  <- gsub(" ","\n",data$dad)
   data$mom  <- gsub(" ","\n",data$mom)}
if(isTRUE(fold)&&!is.na(pairs)){
   pairs$id1 <- gsub(" ","\n",pairs$id1)
   pairs$id2 <- gsub(" ","\n",pairs$id2)
   }

## Generate the pedigree data structure (without no-children couples)
if(is.na(pairs)){ped <- pedigree(id       = data$ID, 
                dadid    = data$dad, 
                momid    = data$mom, 
                sex      = data$sex,
                status   = data$status)} else {

## Generate the pedigree data structure (with no-children couples)
## Create the code (4 for spouse)
pairs$code <- 4
## Process the pedigree
ped <- pedigree(id       = data$ID, 
                dadid    = data$dad, 
                momid    = data$mom, 
                sex      = data$sex,
                status   = data$status,
                relation = pairs) }

## Plot the pedigree chart
plot(ped,
     cex        = cex,
     symbolsize = symbolsize,
     col        = data$color,
     density    = c(-1, -1, -1, -1), ## makes all of the sections solid!
     affected   = data$fill)

  } ## End function ez_ped

## Basic data
data <- read_csv(col_names = TRUE, file=
    "ID,                 dad,           mom,         sex,    status, color,      fill
     William Jones,      NA,            NA,          male,   1,      black,      NA     
     Mary Jones,         NA,            NA,          female, 1,      blue,       1      
     Frank Jones,        William Jones, Mary Jones,  male,   0,      black,      0     
     John Jones,         William Jones, Mary Jones,  male,   1,      black,      0   
     Peter Jones,        William Jones, Mary Jones,  male,   0,      olivedrab,  1     
     Carol Jones,        NA,            NA,          female, 0,      blue,       1      
     LaVerne Jones,      Peter Jones,   Carol Jones, female, 0,      black,      0
     Karen Brown,        NA,            NA,          female, 0,      black,      0
     Samuel Jones-Brown, Frank Jones,   Karen Brown, male,   0,      olivedrab,  1
     Susan Hanford,      NA,            NA,          female, 0,      black,      NA")

## Basic data
data2 <- read_csv(col_names = TRUE, file=
    "ID,    dad,  mom, sex 
     One,   NA,   NA,  male   
     Two,   NA,   NA,  female      
     Three, One,  Two, male     
     Four,  One,  Two, male      
     Five,  One,  Two, male      
     Six,   NA,   NA,  female      
     Seven, Five, Six, female
     Eight, NA,   NA,  female")

## Relationships without children
pairs <- read_csv(col_names = TRUE, file=
                     "id1,        id2
                      John Jones, Susan Hanford")

pairs2 <- read_csv(col_names = TRUE, file=
                     "id1,        id2
                      Eight,      Four")

ez_ped(data, pairs=pairs, symbolsize=2, fold=TRUE)

```

## Bridges

```{r bridges}

data <- read_csv(col_names = TRUE, file=
    "ID,                     dad,                    mom,                     sex,    born, died
    Charles Wentworth,       NA,                     NA,                      male,   1648, 1780
    Amariah Wentworth,       Charles Wentworth,      Bethiah Fenno,           male,   1719, 1799
    Bethiah Fenno,           NA,                     NA,                      female, 1692, 1780      
    Rebecca Shepard,         NA,                     NA,                      female, 1723, 1791
    Elijah Wentworth,        Amariah Wentworth,      Rebecca Shepard,         male,   1775, 1810
    Elijah S. Estes,         NA,                     NA,                      male,   1814, 1887
    Zebiah Walker Wentworth, Elijah Wentworth,       Lucy Walker,             female, 1810, 1887
    Lucy Walker,             NA,                     NA,                      female, NA,   NA
    Mary Adeline Palmer,     NA,                     NA,                      female, 1848, 1918
    Lorenzo Langston Estes,  Elijah S. Estes,        Zebiah Walker Wentworth, male,   1839, 1909
    Addie Estes,             Lorenzo Langston Estes, Mary Adeline Palmer,     female, NA,   NA
    Charles A. Bridges,      NA,                     NA,                      male,   1877, NA
    Charles L. Bridges,      Charles A. Bridges,     Addie Estes,             male,   1912, 2001
    Mary Jane Bridges,       Burr Anderson,          Lulu Anderson,           female, 1912, 2016
    Burr Anderson,           NA,                     NA,                      male,   NA,   NA
    Lulu Anderson,           NA,                     NA,                      female, NA,   NA
    Thomas Bridges,          Charles L. Bridges,     Mary Jane Bridges,       male,   1938, NA
    Kent Bridges,            Charles L. Bridges,     Mary Jane Bridges,       male,   1941, NA")

flextable(data) %>% 
  autofit()
    
ez_ped(data, symbolsize=1, cex= 0.4, fold=FALSE)
```

